### 2020-02-20 Common/chapters/literature_review.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.  Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions. False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate. However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks. In fact, the latter can be considered as single-step CADs. For this reason, CAD systems are currently a trending topic in deep learning.\\
There mainly exists two types of CAD systems based on convolutional neural networks. The first one is called CADe for "computer-aided detection". Its primary goal is "to increase the detection rate of diseases while reducing the false negative  rate  possibly  due  to  the  observers’  mistakes  or  fatigue"~\cite{41}. This group includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection. The other type is CADx for "computer-aided diagnosis" which aims at extracting the characteristics lesions.\\
This chapter goes through the main articles about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on articles that used the same datasets as the ones used in this work.

\section{Prostate/PROSTATEx}
As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally. It was predicted that there would be 1.7 million new cancer cases by 2030. The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"~\cite{41}. Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection. All the articles mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see section \ref{sec:prostatex_dataset_description}).

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65x65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20 to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1. Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist. 
Their model is a modified version of the well-known VGG16 model, including the addition of $1x1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set. The best results were obtained by using DWI images with the highest b-value only, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.

Saifeng et al.~\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$. The AUC on their validation set reached $0.92$. Regarding data processing, their approach stacked different combinations of the available sequences as the three channels instead of defining a single combination like Song et al.: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W. The data augmentation process differs in that the images are rotated in 3D, each lesion being sliced at 7 different orientations. These 2-dimensional slices were then augmented using rotation, shearing and translation of 1px, resulting in 207144 training samples. Both validation and testing test were also augmented in the same manner. This whole process allows to include 3-dimensional information in 2-dimensional images. The method used ensemble learning which combined different models the reach the best performance possible.

Mehrtash et al.~\cite{01} used a different approach. First of all, the input was feeded to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans. Then, each of these feature extractors' outputs were merged into a common decision maker. Furthermore, 3-dimensional convolutions instead 2-dimensional ones were performed. In fact, 3-dimensional patches centered on the lesion were cropped. Augmentation including translation and flipping was used in order to balance the dataset. Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist compared to the previous papers. These tricks allowed their model achieved an AUC of $0.80$ on the PROSTATEx challenge. To make predictions, five different models were used, averaging the predictions of the four best models.

@@

\section{Lung/Lung CT Challenge}
Gao et al. claimed that "lung cancer is one of the most frequent and leading causes to death all over the world. It was reported that there were approximately $1.8x10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).

@@

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/LUNGx_challenge_all_results.png}
\caption{AUC values for the 11 computerized methods and six radiologists in the task of classifying malignant and benign nodules~\cite{12}}
\label{fig:LUNGx_challenge_all_results.png}
\end{figure}

\section{Brain/Kaggle Brain}
Brain cancer is another major type of cancer. According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women. It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}. The following articles are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website. However, since it is the only dataset available for brain tumor classification, some publications used it. Further details about this dataset can be found in section \ref{sec:kaggleBrain}.

Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. Their processing method used a cropping technique which removed extra black margin around the skull. Each border of the image merged with a part of the skull. Since the images come from different sources, their resolution vary quite a lot. Therefore, the authors resized them to $224x224x3$. Moreover, data was augmented by rotation, vertical shifting and horizontal shifting. As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting. The data was split into a training set, a validation set and a test set. Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance. The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$). The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).

>>>>

Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted of two different parts: "Lesion  detection  and  false-positive  reduction.  Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions. False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate. However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks. In fact, the latter can be considered as single-step CADs. For this reason, CAD systems are currently a trending topic in deep learning.\\
There mainly exists two types of CAD systems based on convolutional neural networks. The first one is called CADe for "computer-aided detection". Its primary goal is "to increase the detection rate of diseases while reducing the false negative  rate  possibly  due  to  the  observers’  mistakes  or  fatigue"~\cite{41}. This group includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection. The other type is CADx for "computer-aided diagnosis" which aims at extracting the characteristics lesions.\\
This chapter presents the main research works about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.

\section{Prostate/PROSTATEx}
As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally. It was predicted that there would be 1.7 million new cancer cases by 2030. The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}. Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection. All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1. Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist. 
Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set. The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.

Liu et al.~\cite{46} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$. The AUC on their validation set reached $0.92$. Regarding data processing, their approach stacked different combinations of the available sequences as the three channels instead of defining a single combination like Song et al.: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W. The data augmentation process differs in that the images are rotated in 3D, each lesion being sliced at 7 different orientations. These 2-dimensional slices were then augmented using rotation, shearing and translation of 1px, resulting in 207144 training samples. Both validation and testing test were also augmented in the same manner. This whole process allows to include 3-dimensional information in 2-dimensional images. The method used ensemble learning which combined different models to reach the best performance possible.

Mehrtash et al.~\cite{01} used a different approach. First of all, the input was fed to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans. Then, each of these feature extractors' outputs were merged into a common decision maker. Furthermore, 3-dimensional convolutions instead of 2-dimensional ones were performed. In fact, 3-dimensional patches centered on the lesion were cropped. Augmentation including translation and flipping was used in order to balance the dataset. Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist compared to the previous papers. These tricks allowed their model to achieve an AUC of $0.80$ on the PROSTATEx challenge. To make predictions, five different models were used, averaging the predictions of the four best models.

@@

\section{Lung/Lung CT Challenge}
Gao et al. claimed that "lung cancer is one of the most frequent and leading causes of death all over the world. It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).

@@

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/LUNGx_challenge_all_results.png}
\caption{AUC values for the 11 computerized methods and six radiologists in the task of classifying malignant and benign nodules~\cite{12}}
\label{fig:LUNGx_challenge_all_results}
\end{figure}

\section{Brain/Kaggle Brain}
Brain cancer is another major type of cancer. According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women. It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}. The following research papers are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}. However, since it is the only dataset available for brain tumor classification, some publications used it. Further details about this dataset can be found in section \ref{sec:kaggleBrain}.

Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. Their processing method used a cropping technique which removed extra black margin around the skull. Each border of the image merged with a part of the skull. Since the images come from different sources, their resolution vary quite a lot. Therefore, the authors resized them to $224$x$224$x$3$. Moreover, data was augmented by rotation, vertical shifting and horizontal shifting. As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting. The data was split into a training set, a validation set and a test set. Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance. The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$). The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).

### 2020-02-20 Common/chapters/medical.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

\chapter{Medical information}
\label{ch:medical}
This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly. First of all, some notions about cancer are described in order to understand what it is and which effects it has on the human body. Second, medical data has its own file formats. To make use of them in a deep learning project, medical files must be processed in a certain way, depending on each format. In fact, formats represent 2D, 3D or even 4D data. Some of them require specific normalization in order to get the right rendering. Finally, some visualization tools were developped to display raw medical files easily.

@@

The term cancer refers to different phenomena which involve mutation, abnormal multiplication and spreading of cells. As stated by Hanahan et al. in "The Hallmarks of Cancer" \cite{19} and "The Hallmarks of Cancer: The Next Generation" \cite{20}, every malignant tumor acquires six different capabilities during its evolution: 
\begin{itemize}
	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells don't wait for the body's approval to grow and proliferate, contrary to normal cells. They become the only responsible for their multiplication.
	\item \textbf{"Evading growth suppressors"}\\
The body sends signals to contain cell growth within a tissue. Cancerous cells are insensitive to these. 
	\item \textbf{"Activating invasion and metastasis"}\\
Metastases are cells whose role is to propagate to other parts of the body in order to colonize and create new tumors. 
	\item \textbf{"Enabling replicative immortality"}\\
Healthy cells replication is limited to a certain amount, which is not the case for cancerous cells. 
	\item \textbf{"Inducing angiogenesis"}
\\Angiogenesis is the process of creating new blood vessels. Tumors have an influence on angiogenesis around them, since they need vascularization to continue growing. 
	\item \textbf{"Resisting cell death"}
\\Apoptosis is the programmed death of cells, which is part of the continuous regeneration of every cell within a body. Cancerous cells survive this programmed death.

@@

MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the outputed images as a volume. Different settings, usually called sequences, make the look of the output vary, as shown on figure \ref{fig:PROSTATEx-t2-adc-dwi}.
Unlike MRI, CT is based on X-rays instead of magnetic fields, but still provides a three-dimensional representation of a body part. Figure \ref{fig:visualize_lung_dcm} shows a lung CT scan.

@@

If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
\begin{itemize}
	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
	\item Samples per pixel: the number of channels for each pixel

@@

\subsubsection{Overview of the header structure}
With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes. “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”. Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}

@@

\section{Visualization tools}
Processing data manually increases the probability of making mistakes. For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developped. Their primary goal is to compare our visual representation of an image to the one obtained in professional pieces of software. These tools are convenient to visualize a dataset easily. In some cases (especially RAW/MHD images), no free software capable of reading the files was available, which made the corresponding tool useful.

\subsection{DICOM}
Visualizing DICOM files is pretty straightforward since each file represent a single two-dimensional image. However, the way this standard works require a lot of pixel transformation and normalization to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors. Our tools allows to display a single DICOM file as well as a sequence of files if the function is feeded with a directory. Users can then scroll through the Z-axis, displaying the next or previous slice.

@@

\subsection{NIfTI}
As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp. When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed. Furthermore, the three-dimensionality implies that the volume is viewable under three different perspectives. For example, a three-dimensional brain volume can display it from the top of the head to the bottom, from one ear to the other and from the back of the head to the person's face. Therefore, the user can choose a specific axis to navigate through. If no axis is chosen, all three perspectives are shown one after the other.

@@

\subsection{RAW}
Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e. a succession of 2D slices. Scrolling with the mouse goes through slices over a specific axis. Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part under three different perspectives. This tool was particularly useful since no free software capable of handling these files was available. Photography-related programs such as Photoshop can open RAW pictures but not these three-dimensional ones.

>>>>

\chapter{Medical information}
\label{ch:medical}
This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly. First of all, some notions about cancer are described in order to understand what it is and which effects it has on the human body. Second, medical data has its own file formats. To make use of them in a deep learning project, medical files must be processed in a certain way, depending on each format. In fact, formats represent 2D, 3D or even 4D data. Some of them require specific normalization in order to get the right rendering. Finally, some visualization tools were developed to display raw medical files easily.

@@

The term cancer refers to different phenomena which involve mutation, abnormal multiplication and spreading of cells. As stated by Hanahan et al. in "The Hallmarks of Cancer" \cite{19} and "The Hallmarks of Cancer: The Next Generation" \cite{20}, every malignant tumor acquires six different capabilities during its evolution: 
\begin{itemize}
	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells don't wait for the body's approval to grow and proliferate, contrary to normal cells. They become responsible for their own multiplication.
	\item \textbf{"Evading growth suppressors"}\\
The body sends signals to contain cell growth within a tissue. Cancerous cells are insensitive to these. 
	\item \textbf{"Activating invasion and metastasis"}\\
Metastases are cells whose role is to propagate to other parts of the body in order to colonize and create new tumors. 
	\item \textbf{"Enabling replicative immortality"}\\
Healthy cells replication is limited to a certain amount, which is not the case for cancerous cells. 
	\item \textbf{"Inducing angiogenesis"}
\\Angiogenesis is the process of creating new blood vessels. Tumors have an influence on angiogenesis around them, since they need vascularization to continue growing. 
	\item \textbf{"Resisting cell death"}
\\Apoptosis is the programmed death of cells, which is part of the continuous regeneration of every cell within a body. Cancerous cells survive this programmed death.

@@

MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the generated images as a volume. Different settings, usually called sequences, make the look of the output vary, as shown on figure \ref{fig:PROSTATEx-t2-adc-dwi}.
Unlike MRI, CT is based on X-rays instead of magnetic fields, but still provides a three-dimensional representation of a body part. Figure \ref{fig:visualize_lung_dcm} shows a lung CT scan.

@@

If they are missing, a simple conversion is sufficient. The parameters used to convert the data are given by two tags: 
\begin{itemize}
	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
	\item Samples per pixel: the number of channels for each pixel

@@

\subsubsection{Overview of the header structure}
With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes. “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”. Details about the different fields contained in the header can be found here\footnote{\url{https://brainder.org/2012/09/23/the-nifti-file-format/}}.

@@

\section{Visualization tools}
Processing data manually increases the probability of making mistakes. For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developed. Their primary goal is to compare our visual representation of an image to the one obtained in professional pieces of software. These tools are convenient to visualize a dataset easily. In some cases (especially RAW/MHD images), no free software capable of reading the files was available, which made the corresponding tool useful.

\subsection{DICOM}
Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image. However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors. Our tools allows to display a single DICOM file as well as a sequence of files if the function is fed with a directory. Users can then scroll through the Z-axis, displaying the next or previous slice.

@@

\subsection{NIfTI}
As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp. When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed. Furthermore, the three-dimensionality implies that the volume is viewable from three different perspectives. For example, a three-dimensional brain volume can display it from the top of the head to the bottom, from one ear to the other and from the back of the head to the person's face. Therefore, the user can choose a specific axis to navigate through. If no axis is chosen, all three perspectives are shown one after the other.

@@

\subsection{RAW}
Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e. a succession of 2D slices. Scrolling with the mouse goes through slices over a specific axis. Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part from three different perspectives. This tool was particularly useful since no free software capable of handling these files was available. Photography-related programs such as Photoshop can open RAW pictures but not these three-dimensional ones.
